---
title: "An investigation of the success of linking adults with poor personal of their health and self-care to primary medical care"
author: "Arm and Guy"
date: "May 14, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center")
library(tidyverse)
library(broom)
library(survival)
library(survminer)
library(readr)
library(mosaic)
```

```{r}
urlfile="https://raw.githubusercontent.com/gthampak/Arm_Guy_MATH150_Project/main/HELPdata.csv"

HELPdata <- read_csv(url(urlfile))
```

###Title

###Introduction

(move from general to specific)

background info (look at source 1)

motivation?

###Aim/Hypothesis

Our primary goal was to assess the effectiveness of a novel multi-disciplinary clinic for linking patients in a residential detoxification program with poor personal perception towards healthcare to primary medical care.

###Methods

(make sure replicable)

Explanation of variables we explored, why we explored them (in relation to our mainhypothsis/primary aim)

Categorize variables explored (sf scores??)

###Results (Model goes here!)

Re-state question/goal

Results/Data Analysis of each question/hypothesis

Figures!!

(need connection between ideas)

###Discussion

Does results answer question?

How did we deviate from question in the process? Anything interesting 

Be clear with why we accept of reject null hypotheses

Relate work to previous research

###Sources/References

Perception of Health and Use of Health Care Services in a Swedish Primary Care District. A ten Year's Perspective
https://www.tandfonline.com/doi/pdf/10.3109/02813439109026592









