---
title: "Final_Model"
author: "Arm and Guy"
date: "May 14, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r global_options, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center")
library(tidyverse)
library(broom)
library(survival)
library(survminer)
library(readr)
library(mosaic)
library(pec)
```

```{r}
urlfile="https://raw.githubusercontent.com/gthampak/Arm_Guy_MATH150_Project/main/HELPdata.csv"

HELPdata <- read_csv(url(urlfile))
```

###Treatment Variable

`group` Randomization Group (0=Control, 1=Clinic)

###Variables associated with pain

`b7` Amount of bodily pain -- past 4 weeks (1=None, 2=Very mild, 3= Mild, 4=Moderate, 5= Severe, 6= Very severe)

`b8` Amount of pain interfering with normal work-last 4 weeks (1=Not at all, 2=A little bit, 3=Moderately, 4=Quite a bit, 5=Extremely)

`bp` SF-36 pain index (0-100)

`c2h1` Ever had abdom pain req overnt hosp stay (0=No, 1=Yes)

`c2h2` Abdom pain req ovrnt hosp stay-last 6 months (0=No, 1=Yes)

`c2m1` Chest pain using cocaine req ER/hosp (0=No, 1=Yes)

`c2m2` Chest pain using coc req ER/hosp-last 6 months (0=No, 1=Yes)

`c2o1` Lower back pain > 3 months req med attn (0=No, 1=Yes)

`c2o2` Lwr back pain >3 months req med attention-last 6 months (0=No, 1=Yes)

`rawbp` Raw SF-36 pain index

```{r}
HELPdata_survfit <- survfit(Surv(dayslink, linkstatus) ~ group, data=HELPdata)

ggsurvplot(HELPdata_survfit, conf.type = "TRUE") + 
  ggtitle("KM-curve for Treatment vs. No Treatment")

coxph(Surv(dayslink, linkstatus) ~ group, data=HELPdata) %>% tidy()
```

```{r}

HELPdata_survfit <- survfit(Surv(dayslink, linkstatus) ~ group + hs_grad + alcohol, data=HELPdata)

ggsurvplot(HELPdata_survfit, conf.type = "TRUE")

coxph(Surv(dayslink, linkstatus) ~ group + hs_grad + alcohol, data=HELPdata) %>% tidy()
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + a9, data=HELPdata) %>% tidy()
```


```{r}

HELPdata_survfit <- survfit(Surv(dayslink, linkstatus) ~ group + any_ins, data=HELPdata)

ggsurvplot(HELPdata_survfit, conf.type = "TRUE")

coxph(Surv(dayslink, linkstatus) ~ group, data=HELPdata) %>% tidy()
```


```{r}
HELPdata_survfit <- survfit(Surv(dayslink, linkstatus) ~ group + gh, data=HELPdata)

ggsurvplot(HELPdata_survfit, conf.type = "TRUE")

coxph(Surv(dayslink, linkstatus) ~ group + gh, data=HELPdata) %>% tidy()
```


###Final

Variables

ANY_UTIL Any recent health utilization (0=No, 1=Yes)a
CES_D CES-D score, measure of depressive symptoms, high scores are worse
CNTRL InDUC-2L-Control score
C_AU ASI-Composite score for alcohol use
C_DU ASI-Composite score for drug use
C_MS ASI-Composite medical status

D5_REC Medical trtmt is important y/n (0=No, 1=Yes)
D5 How import is trtmnt for these med probs (0=Not at all, 1=Slightly, 2= Moderately, 3= Considerably, 4= Extremely
E10A have you been to med clinic-last 6 months (0=No, 1=Yes)
E15C2 I did not know where to go for help (0=No, 1=Yes)
E8A1 Saw MD/H care worker regarding alcohol/drugs-last 6 months (0=No, 1=Yes)
FRML_SAT Formal substance abuse treatment y/n (0=No, 1=Yes)

GH SF-36 general health perceptions (0-100)
HT Raw SF-36 health transition item
MH SF-36 mental health index (0-100)
PF SF-36 physical functioning (0-100)
RE SF-36 role-emotional (0-100)
RP SF-36 role physical (0-100)

H14 According to interviewer, which substance is main problem (0=No problem, 1=Alcohol, 2=Alcohol to intox, 3=Heroin 4=Methadone, 5=Other opiate/analg, 6=Barbituates, 7=Sed/hyp/tranq, 8=Cocaine, 9=Amphetamines, 10=Marij/cannabis, 15=Alcohol and one or more drug, 16=More than one drug)
PRIM_SUB First drug of choice (0=None, 1=Alcohol, 2=Cocaine, 3=Heroin, 4=Barbituates, 5=Benzos, 6=Marijuana, 7=Methadone, 8=Opiates)

H19A How import is treatment for alcohol problems now (0=Not at all, 1=Slightly, 2=Moderately, 3=Considerably, 4=Extremely)
H19B How important is trtmnt for drug probs now (0=Not at all, 1=Slightly, 2=Moderately, 3=Considerably, 4=Extremely)

S1A At interview pt obviously depressed/withdrawn (0=No, 1=Yes)
S1B at interview pt obviously hostile (0=No, 1=Yes)
S1C At interview patientt obviously anxious/nervous (0=No, 1

A9 Years of education completed

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + a9 + any_util + ces_d + cntrl + c_au + c_du + c_ms + gh + ht + mh + pf + re + rp + h14 + prim_sub + h19a + h19b + s1a + s1b + s1c, data=HELPdata) %>% tidy()
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + any_util + ces_d + cntrl + c_au + c_du + c_ms, data=HELPdata) %>% tidy()
```


```{r}
coxph(Surv(dayslink, linkstatus) ~ group + d5_rec + d5 + e10a + e15c2 + e8a1 + frml_sat) %>% tidy()
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + gh + ht + mh + pf + re + rp, data=HELPdata) %>% tidy()
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + h14 + prim_sub, data=HELPdata) %>% tidy()
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + h19a + h19b , data=HELPdata) %>% tidy()
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + s1a + s1b + s1c, data=HELPdata) %>% tidy()
```

###SF scores - Model Building

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + pf + rp, data=HELPdata) %>% tidy()
```

###Model Making

Potentially Significant Variables

hs_grad

a9



```{r}
HELPdata_survfit <- survfit(Surv(dayslink, linkstatus) ~ group + hs_grad, data=HELPdata)

ggsurvplot(HELPdata_survfit, conf.type = "TRUE")
```


```{r}
coxph(Surv(dayslink, linkstatus) ~ group + as.factor(h14), data=HELPdata) %>% tidy()
```

Test for hs_grad and substances
```{r}
ggplot(data = HELPdata) +
  geom_bar(aes(x = hs_grad, fill = as.factor(alcohol)))

t.test(any_util~hs_grad, data = HELPdata)
```

```{r}
#v <- c(HELPdata$gh, HELPdata$ht, HELPdata$mh, HELPdata$pf, HELPdata$re, HELPdata$rp)
t.test(d5_rec~alcohol, data = HELPdata)
```


```{r}
HELPdata <- HELPdata %>%
  mutate(pf_grp = case_when(
    pf <= 80 ~ "0-80",
    pf <= 100 ~ "80-100"
  ))
```

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + pf_grp, data=HELPdata) %>% tidy()
```

```{r}
HELPdata_survfit <- survfit(Surv(dayslink, linkstatus) ~ pf_grp, data=HELPdata)

ggsurvplot(HELPdata_survfit, conf.type = "TRUE")
```















