---
title: "An Investigation of How Personal Perception of One's Health and Their Perception of Healthcare affects the effectiveness of Treatments to Link Patients To Primary Medical Care."
author: "Arm and Guy"
date: "May 14, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r global_options, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center")
library(tidyverse)
library(broom)
library(survival)
library(survminer)
library(readr)
library(mosaic)
```

```{r}
urlfile="https://raw.githubusercontent.com/gthampak/Arm_Guy_MATH150_Project/main/HELPdata.csv"

HELPdata <- read_csv(url(urlfile))


```

## Introduction

Connecting with medical care is a choice.  Many different factors could be involved when deciding whether or not to connect with medical care.  Perception of your own personal health influences your decision to connect with medical care for non-major medical emergencies.  Although we are able to do some basic self diagnosis with help of the internet, there is a high chance of a misdiagnosis given the various issues with access to useful and accurate medical information.  Since we do not deem it a necessity if our health is not in a critical condition, we may decide not to connect with medical care.  Apart from the personal health perceptions, there are also perceptions of health care system, which could include the ability to pay for health treatment.  Lastly, one may not be in the optimal mental space to decide whether or not to connect with medical care, which could be due to poor mental health or substance abuse.  

There have been many studies and surveys which try to gauge the accuracy of an individual personal health perception and their views towards healthcare.  We attempt to widen the scope of personal perception of general health to investigate factors which could influence an individual’s view and opinions on their own health as well as primary healthcare.  Better understanding the influences on one's decision to connect with medical care can help focus efforts in specific sectors to give care to people who require care but don't know it themselves.

(move from general to specific)

background info (look at source 1)

motivation?

## Aim/Hypothesis

Our primary goal was to assess how personal perception of someone's health and their perception of healthcare affects the effectiveness of  novel multi-disciplinary clinic for linking patients in a residential detoxification program to primary medical care.

## Methods

First, we looked at variables that are related to health and health perception and separated them into different categories.

* ANY_UTIL Any recent health utilization (0=No, 1=Yes)
* CES_D CES-D score, measure of depressive symptoms, high scores are worse
* CNTRL InDUC-2L-Control score
* C_AU ASI-Composite score for alcohol use
* C_DU ASI-Composite score for drug use
* C_MS ASI-Composite medical status

**Questions related to opinion on and habits towards healthcare**

* D5_REC Medical trtmt is important y/n (0=No, 1=Yes)
* D5 How import is trtmnt for these med probs (0=Not at all, 1=Slightly, 2= Moderately, 3= Considerably, 4= Extremely
* E10A have you been to med clinic-last 6 months (0=No, 1=Yes)
* E15C2 I did not know where to go for help (0=No, 1=Yes)
* E8A1 Saw MD/H care worker regarding alcohol/drugs-last 6 months (0=No, 1=Yes)
* FRML_SAT Formal substance abuse treatment y/n (0=No, 1=Yes)

**SF-36 Scores**

* GH SF-36 general health perceptions (0-100)
* HT Raw SF-36 health transition item
* MH SF-36 mental health index (0-100)
* PF SF-36 physical functioning (0-100)
* RE SF-36 role-emotional (0-100)
* RP SF-36 role physical (0-100)

**Drugs-related variables**

* H14 According to interviewer, which substance is main problem (0=No problem, 1=Alcohol, 2=Alcohol to intox, 3=Heroin 4=Methadone, 5=Other opiate/analg, 6=Barbituates, 7=Sed/hyp/tranq, 8=Cocaine, 9=Amphetamines, 10=Marij/cannabis, 15=Alcohol and one or more drug, 16=More than one drug)
* PRIM_SUB First drug of choice (0=None, 1=Alcohol, 2=Cocaine, 3=Heroin, 4=Barbituates, 5=Benzos, 6=Marijuana, 7=Methadone, 8=Opiates)

**Drug and Healthcare related variables**

* H19A How import is treatment for alcohol problems now (0=Not at all, 1=Slightly, 2=Moderately, 3=Considerably, 4=Extremely)
* H19B How important is trtmnt for drug probs now (0=Not at all, 1=Slightly, 2=Moderately, 3=Considerably, 4=Extremely)

**Interview's perspective on Patient**

* S1A At interview pt obviously depressed/withdrawn (0=No, 1=Yes)
* S1B at interview pt obviously hostile (0=No, 1=Yes)
* S1C At interview patientt obviously anxious/nervous (0=No, 1

**Education-related variables**

* HS_GRAD High school graduate (0=No, 1=Yes)
* A9 Years of education completed

**Model Building**

The functions we used for our primary data analyses are coxph(Surv()) to test for significance of Hazard Ratio coefficients and survfit(Sruv()) to plot survival curves. First we put all the variables above into a single model and removed the ones with highest significance one at a time. We also ran models with variables from a single category (above), and removed variables with highest signifiance. We did this to prevent putting highly correlated variables from the same category into the main model.

(make sure replicable)

Explanation of variables we explored, why we explored them (in relation to our mainhypothsis/primary aim)

Categorize variables explored (sf scores??)

## Results (Model goes here!)

**Significant Variables after prelimiinary models**
* group
* alcohol
* coc_her
* hs_grad
* a9 (education)
* pf
* any_util

**First Actual Model**

```{r}
coxph(Surv(dayslink, linkstatus) ~ group + alcohol + coc_her + hs_grad + pf + any_util, data=HELPdata) %>% tidy()
```

Re-state question/goal

Results/Data Analysis of each question/hypothesis

Figures!!

(need connection between ideas)

## Discussion

Does results answer question?

How did we deviate from question in the process? Anything interesting 

Be clear with why we accept of reject null hypotheses

Relate work to previous research


**Interesting points (potentially)**
* Variables that are significant tend to be binary variables with an even distribution of yes/no answers (or even more lob-sided towards yes (1) response). We suspect this is because more observations in each group directly translates to higher power, a lower absolute difference can result in lower (and potentially significant) p-values.


## Sources/References

Perception of Health and Use of Health Care Services in a Swedish Primary Care District. A ten Year's Perspective
https://www.tandfonline.com/doi/pdf/10.3109/02813439109026592









