---
title: "An investigation of the success of linking adults with poor personal of their health and self-care to primary medical care"
author: "Arm and Guy"
date: "May 14, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center")
library(tidyverse)
library(broom)
library(survival)
library(survminer)
library(readr)
library(mosaic)
```

```{r}
urlfile="https://raw.githubusercontent.com/gthampak/Arm_Guy_MATH150_Project/main/HELPdata.csv"

HELPdata <- read_csv(url(urlfile))
```

###Title

###Introduction

Connecting with medical care is a choice.  Many different factors could be involved when deciding whether or not to connect with medical care.  Perception of your own personal health influences your decision to connect with medical care for non-major medical emergencies.  Although we are able to do some basic self diagnosis with help of the internet, there is a high chance of a misdiagnosis given the various issues with access to useful and accurate medical information.  Since we do not deem it a necessity if our health is not in a critical condition, we may decide not to connect with medical care.  Apart from the personal health perceptions, there are also perceptions of health care system, which could include the ability to pay for health treatment.  Lastly, one may not be in the optimal mental space to decide whether or not to connect with medical care, which could be due to poor mental health or substance abuse.  

There have been many studies and surveys which try to gauge the accuracy of an individual personal health perception and their views towards healthcare.  We attempt to widen the scope of personal perception of general health to investigate factors which could influence an individual’s view and opinions on their own health as well as primary healthcare.  Better understanding the influences on one's decision to connect with medical care can help focus efforts in specific sectors to give care to people who require care but don't know it themselves.
(move from general to specific)

background info (look at source 1)

motivation?

###Aim/Hypothesis

Our primary goal was to assess the effectiveness of a novel multi-disciplinary clinic for linking patients in a residential detoxification program with poor personal perception towards healthcare to primary medical care.

###Methods

(make sure replicable)

Explanation of variables we explored, why we explored them (in relation to our mainhypothsis/primary aim)

Categorize variables explored (sf scores??)

###Results (Model goes here!)

Re-state question/goal

Results/Data Analysis of each question/hypothesis

Figures!!

(need connection between ideas)

###Discussion

Does results answer question?

How did we deviate from question in the process? Anything interesting 

Be clear with why we accept of reject null hypotheses

Relate work to previous research

###Sources/References

Perception of Health and Use of Health Care Services in a Swedish Primary Care District. A ten Year's Perspective
https://www.tandfonline.com/doi/pdf/10.3109/02813439109026592









